Temporality is Important for Networks of Hospital Patient Transfers

### 2025-04-08 reports/report.qmd
small edit: small analytical example

# Methods

@@

A weighted, directed network can be used to construct a Markov chain that in turn can be analysed or simulated to understand

@@

>>>>

# Models and Methods

@@

A weighted, directed network can be used to construct a Markov chain that in turn can be analysed or simulated to understand the dynamics of a spreading infectious disease.

@@

These models have distinct advantages and disadvantages: the "simpler" models are more amenable to analysis, but the more "complex" models represent more fidelity.
For example, the snapshot network model allows us to analyse each static snapshot as its own graph, and then subsequently aggregate those results as a time series; this is a much smaller algorithmic complexity compared to any analogous analysis that could be done on the full temporal network.

In the extreme case, we can analyse a toy model in order to demonstrate the difference between the naïve and improved static network models.

Consider a two-node network, where patients transfer from hospital A to hospital B. Suppose over some time period $T$ there are $n_d$ direct transfers and $n_i$ indirect transfers.
In the naïve model, the expected hitting time is equal to $\frac{T}{n_d+ n_i}$.
In the improved model, we suppose we also have the mean inter-admission time of $\tau$.
Then the expected hitting times $k$ can be solved by a linear system [@Norris_Markov_Chains_1997]:

$$
\begin{bmatrix}
\frac{T}{n_d+n_i} & -\frac{T}{n_d} & -\frac{T}{n_i} \\
0 & 1 & 0 \\
0 & -\frac{1}{\tau} & \frac{1}{\tau}
\end{bmatrix} \begin{bmatrix} k_A \\ k_B \\ k_H \end{bmatrix}
= \begin{bmatrix} 1 \\ 0 \\ 1\end{bmatrix}
$$

which solves to give $k_A = \frac{T}{n_d+n_i}\left(1 + \frac{n_i}{T} \tau \right)$.
Somewhat expectedly, the expectation of the time taken to reach hospital B from hospital A is longer by a factor related to the number of indirect movements and their average inter-admission period.

# Simulation Experiments

Beyond simple models, there is vanishingly small possibility of finding tractable analystical expressions for the hitting times.

## Experimental Setup

<hr />

# APDX
